\large{Vipassana for Hackers} \\ \Huge{The Proposal} \\ \large\textit{Version 0.1}

### 2024-06-26 the-proposal/the-proposal.tex
long proposal: citations for intro

\begin{abstract}
Vipassana meditation (as taught by S.N. Goenka in the tradition of Sayagyi U Ba Khin)
is unlike other meditation techniques insofar as its claims to completeness and
outcomes are concerned. Vipassana claims to ultimately explore the entire field of
mind and matter, with the goals of total liberation and full enlightenment. Implicit
within these claims is a complete understanding of human consciousness. These are
difficult claims to prove because the time commitment required to research the
technique orders on multiple decades --- if not multiple generations. This is complicated by the
fact that the time commitment is demanded of both the researcher \textit{and the
subject}. Until now, research on meditation of all kinds has covered only one of
two fields: (1) controlled experiments which must inherently rely on superficial data
gathered from beginners and (2) observational exploratory research of monks --- expert
meditators who have dedicated their lives to the practice. I propose bridging this
gap by submitting myself to the middle ground. While remaining a layperson, I will
commit to a high ratio of waking medition hours for the rest of my life. Individually
and internally, I will conduct qualitative research into the consequences of
Vipassana meditation and the nature of consciousness while collectively and
externally pledging myself as a subject for long-term quantitative
studies with a broader community of researchers.

\todo{add references to (1) and (2)}

>>>>

\begin{abstract}
Vipassana meditation claims to ultimately explore the entire field of
mind and matter, with the goal of total liberation from suffering. Implicit
within these claims is a complete understanding of human consciousness. These are
difficult claims to prove or disprove because the time commitment required to research the
technique orders on multiple decades --- if not multiple generations.
This is complicated by the
fact that the time commitment is demanded of both the researcher \textit{and the
subject}. Until now, research on meditation of all kinds has covered only one of
two fields:
(1) controlled experiments which must inherently rely on superficial data
gathered from beginners \cite{https://doi.org/10.1002/jclp.20761,
  https://doi.org/10.1037/a0018441,
  https://doi.org/10.1093/scan/nss054,
  https://doi.org/10.31231/osf.io/hzv65}
and (2) observational exploratory research of monks --- expert
meditators who have dedicated their lives to the practice.
\cite{https://doi.org/10.1073/pnas.0407401101,
https://doi.org/10.1371/journal.pone.0073417}

I propose bridging this
gap by submitting myself to the middle ground. While remaining a layperson, I will
commit to a high ratio of waking medition hours. Individually
and internally, I will conduct qualitative research into the consequences of
Vipassana meditation and the nature of consciousness while collectively and
externally pledging myself as a subject for long-term quantitative
studies with a broader community of researchers.
